---
title: "Energy Sector Weekly Report V1.1"
author: "[François de Ryckel](https://fderyckel.github.io/aboutme/)<br>I'm on [Twitter](https://twitter.com/Read_The_Dung) and on [Github](https://github.com/fderyckel)"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: 
  html_document:
    highlight: tango
    theme: spacelab
    toc_float: true
    toc: yes
    toc_depth: 2
    df_print: paged
---

This is the weekly **Energy Sector** financial summary.  

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, 
  warning = FALSE, 
  message = FALSE
)


```

```{r load_libraries}
library(tidyverse)
library(forcats)
library(lubridate)
library(scales)
library(bdscale)

# setting up the path within the project for easier portability of code
thePath <- here::here()

source(paste0(thePath, "/functions/get_data.R"))
source(paste0(thePath, "/functions/augment_data.R"))
source(paste0(thePath, "/functions/visualize_data.R"))

tickers <- readr::read_csv(paste0(thePath, "/10_energy/", "tickers.csv"))
tickers$ticker_stock_exchange <- as_factor(tickers$ticker_stock_exchange)
tickers$ticker_industry <- as_factor(tickers$ticker_industry)
tickers$ticker_sector <- as_factor(tickers$ticker_sector)
tickers <- tickers %>% arrange(ticker_industry, ticker_sector, ticker_subsector)

#tickers <- tickers[52:66, ] 
```

```{r downloading_data, eval = FALSE}
#get_eod_data(tickers)

#download data otherway
yo <- tickers %>% #filter(api_ticker == "CJ") %>% # filter(en_biggest_perc == "ETF") %>% 
  select(api_ticker, ticker_stock_exchange) %>% 
  mutate(yoo = walk2(api_ticker, ticker_stock_exchange, get_eod_data_II))

```

```{r loading_data}
tickerss <- tickers[1:30, ] %>% select(api_ticker, ticker_stock_exchange, ticker_industry)
#tickerss <- tickers %>% select(api_ticker, ticker_stock_exchange, ticker_industry)
df <- tickerss %>% filter(is.na(api_ticker) == 0) %>% 
  mutate(raw_fin_data = map2(api_ticker, ticker_stock_exchange, load_fin_data), 
         augmented_fin_data = map(raw_fin_data, possibly(create_big_df, otherwise = NA)))

```


The *Weekly Energy Report* attempt to see the big trends in the energy sectors[^0]. 

We are mainly interested in understanding the transition from an **oil world** to an **energy world**. 

At the moment, the report is build around `r nrow(tickers)` companies.  You can see their tickers, names, sectors and sub-sectors at [the end](#list_financial_instru) of the report. We created chart for each of these companies.  The whole bundle of charts available [here](https://fderyckel.github.io/WER/oil_gaz.pdf). 

Also, the focus is really on North-American Oil & Gas companies.  We will address this shortcoming ~~soon~~ by adding more European and Asian energy companies to the mix as well as diversifying the industries to include uranium, electricity generations, renewable energies and industrials[^1].  

[^1]: We use the *AlphaVantage* [API](https://www.alphavantage.co/) to get our data.  

[^0]: The report has been generated with the R language and thanks to the many community members who have contributed their time in building packages.  
 
We would love to hear from you. Are there typos, mistakes?  Are graphs unclear?  not useful? Should we add additional resources and analysis?  Write us a [mail](mailto:f.deryckel@gmail.com)    

To read the last report, go [here](https://fderyckel.github.io/WER/reports/WER_2018_09_03.html) and to see the second to last one go [here](https://fderyckel.github.io/WER/reports/WER_2018_06_11.html)

# What made the news ?  {.tabset}

## Oil and Gas
* Big news are: 
    + 

## Renewable Energies
*  

## Uranium  


## Utilities - Electricity Generations
* 

# Report higlights. 

*  


# General Energy overview {#ener_overview} 

List of ETF selected for the general overview.  
```{r energy_global_etf_choice}
df_ticker <- tibble(ticker = c("KOL", "IXC", "UNG", "MLPX", "URA", "SPY", "VDE"))

df2 <- inner_join(df_ticker %>% select(api_ticker = ticker), tickers, by = "api_ticker") %>% 
  select(ticker = api_ticker, Name = ticker_name, ticker_stock_exchange)

knitr::kable(df2 %>% select(-ticker_stock_exchange), caption = "Selected Energy related ETF")
```

Returns of the energy ETF over the last 3 months
```{r energy_3months_return, out.width = 900}
df3 <- df2 %>% rename(api_ticker = ticker) %>% 
  mutate(price_data = map2(api_ticker, ticker_stock_exchange, load_fin_data), 
         adjusted_return = map(price_data, 
                               function(.x) get_adjusted_return(.x, start_date = today() - 63))) %>% 
  select(-price_data, -ticker_stock_exchange) %>% unnest() 

plot_adjusted_return(df3, "3 months ETF return")
```

Returns of the energy ETF over the last 12 months
```{r energy_12months_return, out.width = 900}
df3 <- df2 %>% rename(api_ticker = ticker) %>% 
  mutate(price_data = map2(api_ticker, ticker_stock_exchange, load_fin_data), 
         adjusted_return = map(price_data, 
                               function(.x) get_adjusted_return(.x, start_date = today() - 365))) %>% 
  select(-price_data, -ticker_stock_exchange) %>% unnest()

plot_adjusted_return(df3, "12 months ETF return")
```



# The Oil & Gas Industry   {.tabset}
Following GISC, the Oil & Gas (OG) industry consists of the following sectors and sub-sectors: 
```{r display_sectors_OG}
yo <- tickers %>% filter(ticker_industry == "oil_gas") %>% arrange(GISC_Code) %>% 
  select(ticker_sector, ticker_subsector, ticker_name) %>% 
  group_by(ticker_sector, ticker_subsector) %>% summarize(numerb = n()) %>% 
  #unique() %>% 
  na.omit() %>% select(Sector = ticker_sector, Subsector = ticker_subsector, `Number of stocks` = numerb)

yo$Sector = str_replace_all(yo$Sector, c("energy_equipment_services" = "Equipment & Services", 
                                         "oil_gas_consumable_fuels" = "Consumable Fuels"))

yo$Subsector = str_replace_all(yo$Subsector, c("drilling" = "Drilling", 
                                               "equipment_services" = "Equipment & Services Cies", 
                                               "exploration_production" = "Exploration & Production Cies", 
                                               "integrated" = "Major Integrated Cies",
                                               "refining_marketing" = "Refining & Marketing Cies", 
                                               "storage_transportation" = "Storage & Transportation Cies"))

knitr::kable(yo, caption = "List of Sectors and Sub-sectors of the O&G Industry")
```

[At the bottom of this file](##list_financial_instru), there is the complete list of financial instruments with their sectors and sub-sectors on which the following analysis has been based on. 

## Long term trends
Let's check the main Oil & Gas ETF returns. 
We have picked the following ETF
```{r og_etf}
og_etf <- c("XOP", "OIH", "VDE", "CHIE")
df2 <- tickers %>% filter(api_ticker %in% og_etf) %>% 
  select(Ticker = api_ticker, Name = ticker_name)
knitr::kable(df2)
```

```{r etf_oneyar_return, echo=TRUE, eval=FALSE, out.width = 900}
df2 <- tibble(ticker = og_etf)

df2 <- df2 %>% 
  mutate(price_data = map(ticker, function(.x) read_fin_data(.x))) %>% 
  unnest()
df2$Index <- ymd(df2$Index)


df2 <- df %>% rename(api_ticker = ticker) %>% 
  mutate(price_data = map2(api_ticker, ticker_stock_exchange, load_fin_data), 
         adjusted_return = map(price_data, 
                               function(.x) get_adjusted_return(.x, start_date = today() - 365))) %>% 
  select(-price_data, -ticker_stock_exchange) %>% unnest()


plot_adjusted_return(df2, "12 months ETF return")

```

## Breadth of the Oil & Gas Market  {#og_breadth}
Usually, on shorter time frame, we use the 50 days Standard Moving Average (SMA) to see if a stock is in an uptrend.  The graphs below does exactly that: it checks the percentage of stocks that are trading above their 50 days SMA.  This also tell us when the sector might be over-extended to the upside or to the downside.  

```{r percent_stock_above_sma50, echo=FALSE, warning=FALSE, out.width = 700, dpi = 300}
df2 <- df %>% select(api_ticker, augmented_fin_data) %>% 
  filter(is.na(augmented_fin_data) == 0) %>% 
  unnest() %>% 
  select(api_ticker, Index, sma50_abov) %>% #filter(Index == today() - 1)
  mutate(sma50_abov_bin = if_else(sma50_abov > 0, 1, 0)) %>% group_by(Index) %>% 
  summarise(number_stocks = n(), 
            stock_above_sma50 = sum(sma50_abov_bin, na.rm = TRUE)) %>% 
  mutate(percentage_above = stock_above_sma50 / number_stocks)

# filter df2 for the last 3 years
df3 <- df2 %>% filter(Index >= today() - 3*365) %>% 
  #filtering out the days some markets are not open
  mutate(market_open = if_else(number_stocks < 0.75 * lag(number_stocks), 0, 1)) %>% 
  filter(market_open == 1)

# Create our breadth indicator graph 
ggplot(df3, aes(x = Index, y = percentage_above)) + 
  geom_line() + 
  geom_hline(yintercept = 0.875, color = "red", linetype = "dashed") + 
  geom_hline(yintercept = 0.125, color = "red", linetype = "dashed") + 
  scale_y_continuous(labels = scales::percent) + 
  scale_x_date(name = "Dates", date_breaks = "4 months", date_labels = "%b %y") + 
  theme(axis.title.y = element_blank(), 
        axis.title.x = element_blank(), 
        axis.text.x = element_text(angle = 60, vjust = 0.5), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.minor.y = element_line(colour = "grey50", linetype = "dashed"), 
        panel.grid.major.y = element_line(colour = "grey80", linetype = "dashed")) + 
  ggtitle("Percentage of stocks above their 50 SMA")
```


## Weekly and bi-monthly returns  {#weekly_returns} 
In this section, we check the weekly return for the last 12 weeks for all stocks belonging to the industry, sectors and sub-sectors.  
```{r main_df_creation, eval = FALSE, echo = TRUE}
df2 <- df %>% select()
  filter(days_of_week == "Monday" & ticker_industry == "oil_gas") %>% group_by(ticker) %>% 
  arrange(desc(Index)) %>% 
  filter(row_number() %in% c(1:12)) %>% 
  mutate(bi_monthly_yes = rep(seq(0, 1), 6)) %>% 
  ungroup()

```


```{r weekly_return, warning=FALSE, echo = FALSE, out.width = 900, dpi = 300}
# Function to detect ouliers-ish (very loose definition)
is_outlier <- function(x) {
  return(x < quantile(x, 0.25) - 0.95 * IQR(x) | x > quantile(x, 0.75) + 0.95 * IQR(x))
}

yo <- ymd(as.character(df$Index)) 
yo <- tibble(dates = yo) %>% group_by(dates) %>% summarise(yup = unique(dates)) %>% arrange(desc(dates)) 

df %>% filter(Index >= yo$dates[12]) %>% 
  mutate(outlier = if_else(is_outlier(weekly_return), ticker, as.character(NA))) %>% 
  ggplot(., aes(x=as.factor(Index), y = weekly_return, col = Index)) + geom_boxplot() + 
  #scale_colour_gradient(low = "dodgerblue1", high = "dodgerblue4") + 
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=50, hjust=.5, vjust=.5), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.minor.y = element_line(colour = "grey50", linetype = "dashed"), 
        panel.grid.major.y = element_line(colour = "grey80", linetype = "dashed"), 
        legend.position = "none") + 
  geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 2, check_overlap = TRUE) + 
  geom_hline(yintercept = 0, color = "red") + 
  ggtitle("Energy - Weekly Return") + ylim(c(-15, 15)) + 
  labs(x= "Dates", y="Weekly Return (in %)") 
```

We do the same graph but with a 2 weeks return and a disaggregation by sub-sector.  
```{r bi_monthly_return, warning=FALSE, echo = FALSE, out.width = 900, dpi = 300}
# bi-monthly return boxplot ordered with sub-sector
n <- seq(0:5) - 1
dates <- tibble(value = yo$dates[1] - 14*n)
df %>% filter(Index %in% unlist(dates)) %>% 
  group_by(Index) %>% 
  #mutate(outlier = if_else(is_outlier(bi_monthly_return), ticker, as.character(NA))) %>% 
  ggplot(., aes(x=as.factor(Index), y = bi_monthly_return, fill = ticker_subsector)) + geom_boxplot() + 
  theme(axis.text.x = element_text(colour="grey20", size=8, angle=50, hjust=.5, vjust=.5), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.minor.y = element_line(colour = "grey50", linetype = "dashed"), 
        panel.grid.major.y = element_line(colour = "grey80", linetype = "dashed"), 
        legend.position = "top", legend.title = element_blank()) + 
  #geom_text(aes(label = outlier), na.rm = TRUE, hjust = -0.3, size = 2, check_overlap = TRUE) + 
  ggtitle("Energy - 2-Weeks Return") + ylim(c(-15, 15)) + 
  labs(x= "Dates", y="Bi-monthly Return (in %)")

```

## Recap table  
```{r trend_table_OG_prep, echo = FALSE, warning=FALSE}
yo <- df2 %>% filter(ticker_industry == "oil_gas") %>% group_by(ticker) %>% 
  filter(Index == max(Index)) %>% 
  ungroup() %>% 
  mutate_at(7:16, round, 2) %>% 
  select(Index, Ticker = ticker, `Ticker Name`= ticker_name, 
         `Rolling std 50 sma`= diff_sd_sma50, 
         `Monthly Sharp R`= monthly_sharp_Ratio, `Yearly Sharp R` = yearly_sharp_Ratio, 
         `Weekly return` = weekly_return, `Bi-monthly return` = bi_monthly_return, 
         `Monthly return` = monthly_return, 
         `Semester return` = semester_return, `Yearly return` = yearly_return)
yu <- yo$Index[1]
```

Now let's have a recap table to drill in at the company level.  This table has been generated with data from `r format(yu, '%d %B, %Y')`
```{r trend_table_OG, echo = TRUE, eval = FALSE}
## Make a js table with all the variable that can be sorted.  
DT::datatable(yo %>% select(-Index), options = list(pageLength = 10)) %>% 
  DT::formatRound(columns = 11, digits = 2)
```


# Uranium {.tabset}

# The list of financial instruments  {#list_financial_instru}
Here is the list of all equities used in this report, their sector and sub-sector.  
```{r equities_table, echo=TRUE, eval=FALSE}
#do graph by industry, sector
#order the bar by decreasing order.  
DT::datatable(tickers %>% select(-yahoo_ticker, -ETF, -GISC_Code), options = list(pageLength = 10), 
             caption = "List of stocks, sectors and sub-sectors") 
```

If you have enjoyed this report, [let us now](mailto:f.deryckel@gmail.com).  

